---
title: "STAT 424 HW1"
subtitle: Fall 2020
output: pdf_document
pdf_document: default
word_document: default
---

## Instructions

1. Submit your solution on Canvas as a pdf file before *September 11 at 11:59pm*.
2. Use R for all computation, and include your code. 
3. We recommend you start with the template `hw1.Rmd` file on Canvas.

## Rubric

* [30 points] One problem below will be graded for,
  - [15 points] Correctness. Are all parts answered correctly?
  - [15 points] Clarity. Are all answers justified? For questions with a coding
  component, is code provided, and is it readable?
* [20 = 4 * 5 points] Four problems below will be graded on completeness.
* Total: 50 points

```{r setup, include=FALSE}
library("tibble")
knitr::opts_chunk$set(echo = TRUE)
```

## Problems
```{r}
library("ggplot2")
```


**1. [DAE 1.3]** Suppose that you want to compare the growth of garden flowers
with different conditions of sunlight, water, fertilizer, and soil conditions.
Complete steps 1 - 3 of the guidelines for designing experiments in Section 1.4.
This is an open-ended question -- there is no one right answer -- just explain
your thought process.

In order to design an experiment for the growth of garden flowers, we must first recognize the problem. In this scenario, we must quantify the affect that sunlight, water, fertilizer, and soil conditions have on the growth of the garden flowers. The problem is discovering each variable's affect on the growth of the flower. The response variable in this problem is the height of the garden flowers. The different factors that affect the height of the flowers are sunlight, water, fertilizer and soil conditions.

**2. [DAE 1.8 / MED 1.5]**

  a. What is replication? Why do we need replication in an
  experiment?
  Replication is the act of running multiple trials for each combination of variables. We need replication to estimate the error and to create a larger sample size so we can become more confident in our results.
  b. Present an example that illustrates the difference between
  replication and repeated measurements.
  Let's say we are measuring the affect of various plants a frog eats on the frog's reflexes. Replication would be the frog's varying response time to different plants that it eats. Repeated measurements would be repeating the same factors on the same frog multiple times.
  c. State whether or not replication in your example would have any
  nonstatistical disadvantages, such as a considerable increase in the cost of
  running the experiment.
  After many trials, we wouldn't just be measuring the affect that one plant has on a frog's reflexes. We would be measuring the cumulative affect of all the plants the frog has digested.
**3. [DAE 1.9, modified]**

  a. Why is randomization important in an experiment? 
  Randomization is important because we want the errors to be independently distributed among the random variables.
  b. Give one example where complete randomization would be difficult or
  unethical^[The field of observational causal inference exists for drawing
  conclusions in this case, but it is out of scope for our class.].
  It may difficult to complete randomization in a trial involving temperature. It can be costly in terms of time and resources to randomly change temperature for an experiment.

**4. [MED, 1.4, modified]** An experiment with a single factor and two levels, 1 and
2, was used, and the results were as follows:

```{r}
df <- tibble(
  "sample_id" = as.factor(seq(10)),
  "1" = c(6.1, 8.2, 7.3, 8.4, 8.0, 7.6, 8.7, 9.3, 6.8, 7.5),
  "2" = c(6.3, 8.0, 7.7, 8.1, 8.6, 7.2, 8.4, 9.7, 6.8, 7.2)
)
```

  a. Make two histograms, one for when factor 1 was used, another for when factor
  2 was used.
```{r}
m1=matrix(data=c(1,2),nrow=2,ncol=1,byrow=TRUE)
qplot(df$"1", geom="histogram", binwidth=0.5, xlab="Factor=1",ylab="count")
qplot(df$"2", geom="histogram", binwidth=0.5, xlab="Factor=2",ylab="count")
```

  
  b. Plot the data with the sample ID as the $x$-axis, the measured value as the
  $y$-axis, and the factor level as the color of the point.
  
```{r}
ggplot(df)+geom_point(aes(x=sample_id,y=`1`,colour=factor("1")), shape="circle")+geom_point(aes(x=sample_id,y=`2`, colour=factor("2")),shape="triangle")+labs(x="sample_id",y="measured value",colour="Factors")
```

  
  c. What does part (b) tell you about the need for blocking? Is it needed -- why
  or why not?
  It tells you that the measured value changes based on the sample id. It is needed because the value changes based on the sample id, not just the factor.
    
**5. [MED 1.15]** Assume that an experiment with four levels of a single factor
was run, and the value of the response variable was almost strictly increasing
during the runs of the experiment. What would you suspect and what would be your
recommendation?

I would expect there to be some variable that is affecting the runs of the experiment. I would recommend blocking for that variable and then rerun the experiment.

## Feedback

a. How much time did you spend on this homework?
Approximately 2 hours
b. Which problem did you find most valuable?
Question 4. I liked the practice with Rstudio.

## References

[DAE] Montgomery, Douglas C. Design and analysis of experiments. John Wiley &
sons, 2017.

[MED] Ryan, Thomas P. Modern experimental design. Hoboken, N.J:
Wiley-Interscience, 2007. Print. Accessed for free at
https://search.library.wisc.edu/catalog/9912216524202121.
